\chapter{Abstract}

Computer Aided Engineering (CAE) systems are a type of software which aggregate numerical solvers for Partial Differential Equations (PDEs), $3$D CAD (Computer Aided Design) software, optimization algorithms and all the other software required for a complete assessment of the performance of a given engineering design and for its optimization. One of their most important field of application is Computational Fluid Dynamic (CFD).

\medskip
Nowadays the solvers integrated in CAE systems are typically based on Finite Element Method (FEM) or Finite Volume Method (FVM): their common ground is their need for domain discretization by means of a mesh. However the process of the mesh creation has several limitations which hinder the solver flexibility and its integration with the other software packages present in CAE systems.
To overcome this limitation it is possible to use different solvers that does not rely on the creation of a mesh for solving the PDEs: these are called Meshless or Meshfree Methods (MMs).

\medskip
The goal of this thesis has been the implementation of an optimizer for one and three-dimensional design optimization problems based on simple models described by means of Poisson equations and solved by the Radial Basis Function-Finite Difference (RBF-FD) meshless method. To do so we studied the Automatic Differentiation (AD) and the adjoint method, and we implemented them on a gradient-based optimization algorithm.

\medskip
To asses the quality of the implemented method we applied it to a generic $1$D optimization problem and to a specific type of $3$D optimization problem. In the latter case we also compared the results obtained by our method to those obtained with the continuous adjoint.
